Cryptanalysis of SBC based post-quantum signature scheme

### 2025-06-27 rapport.tex
intro + cGGM

% !TeX spellcheck = en_US
% !TXS template
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper]{geometry}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[backend=bibtex]{biblatex}
\usepackage{csquotes}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage[most]{tcolorbox}
\usepackage{enumitem}
\usepackage{fancyvrb}
\usepackage{multicol}
\usepackage{stmaryrd}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{caption}
\usetikzlibrary{arrows.meta, positioning}

@@

	\begin{abstract}
		The \emph{MPC-in-the-Head} (MPCitH) paradigm, first introduced in \cite{IKOS07}, has recently gained significant attention—particularly in the second round of NIST’s post-quantum signature standardization process, where 5 out of 14 candidate schemes were based on this approach.

@@

@@

		The objective of this internship is to conduct a cryptanalysis of this scheme, assess its security and efficiency, and explore potential improvements.
	\end{abstract}

	\section{MPCitH and SBC signature scheme}

@@

		\subsubsection{GGM tree}
		Since most of the values used in the MPC protocol are randomly generated, a technique known as the GGM tree can be employed to transmit only a seed, rather than all the individual random values.

@@

@@

	\section{Polynomial system solving}
		Gröbner basis theory was first introduced by Buchberger in his thesis \cite{Buc} and he gave an algorithm to compute the Gröbner basis of an ideal.
		Gröbner basis algorithms like Buchberger's Algorithm \cite{Buc}, are used to find solutions for polynomial systems using algebraic geometry tools. For a gentle introduction to this subject \cite{CLS}.

@@

	\section{Contributions}

@@

		\begin{tcolorbox}[title=Protocol 15: Signing Algorithm, breakable, enhanced, colback=white, colframe=black]\label{Protocol_sig}
			\textbf{Inputs:} A public key $x = (\vec{u}, \vec{v})$, a private key $w = (\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ and a message $m \in \{0,1\}^*$.

@@

			\vspace{0.5em}
			\textbf{Phase 4: Second challenge (parties to be opened).}
			\begin{enumerate}[label=\arabic*.]
				\item Compute $h_2 := \text{Hash}(m, \texttt{salt}, \texttt{comm}^{[1]}, \dots, \texttt{comm}^{[\tau]})$.
				\item Expand $h_2$ as $(I^{[e]})_{e \in [1:\tau]} \leftarrow \text{Expand}(h_2)$ where, for every $e, I^{[e]} \in [1:N]$ the party that will not be opened.
			\end{enumerate}

@@

			$$
			\sigma := \texttt{salt} \parallel
			h_1 \parallel
			punctured\_keys \parallel
			[\delta_A^{[e]}]_{e\in[1,\tau]} \parallel
			[\delta_B^{[e]}]_{e\in[1,\tau]} \parallel
			$$
			$$
			[\vec{\delta}_x^{[e]}]_{e\in[1,\tau]} \parallel [\vec{\delta}_y^{[e]}]_{e\in[1,\tau]} \parallel
			[(X_1 + t_0(\vec{u} \cdot \vec{x}))^{[e]}]_{e\in[1,\tau]} \parallel
			[(X_2 + t_0(\vec{v} \cdot \vec{x}))^{[e]}]_{e\in[1,\tau]} \parallel
			$$
			$$
			[(Y_1 + t_0(\vec{v} \cdot \vec{y}))^{[e]}]_{e\in[1,\tau]} \parallel
			[(Y_2 + t_0(\vec{u} \cdot \vec{y}))^{[e]}]_{e\in[1,\tau]}
			$$
		\end{tcolorbox}

@@

>>>>

% !TeX spellcheck = en_US
% !TXS template
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[a4paper]{geometry}
\usepackage{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage[backend=bibtex,style=alphabetic]{biblatex}
\usepackage{csquotes}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{hyperref}
\usepackage[most]{tcolorbox}
\usepackage{enumitem}
\usepackage{fancyvrb}
\usepackage{multicol}
\usepackage{stmaryrd}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{caption}
\usetikzlibrary{arrows.meta, positioning}

@@

	\section{Introduction}
		Since the emergence of quantum computing in the late 1990s, several breakthroughs have been achieved in the field of quantum algorithms. Some of these advancements have found applications in areas such as medicine and physics, while others have demonstrated the ability to break cryptographic problems that are currently assumed to be hard for classical computers, such as RSA and ECDH.

		In particular, quantum algorithms like Grover's \cite{Grover} and Shor's \cite{Shor} have shown that existing public-key cryptosystems are vulnerable in a post-quantum setting. This has led to an urgent need for new hard problems on which to base quantum-resistant public-key cryptography.

		Post-quantum cryptographic schemes are generally built upon four main problem classes: isogeny-based, multivariate, code-based, and lattice-based cryptography. In recent years, new paradigms have also been explored to improve the efficiency and security of schemes within these classes. One such paradigm is the \emph{MPC-in-the-Head} approach.

		Although these underlying problems are believed to be hard even for quantum computers, some have been shown to be vulnerable to classical attacks due to hidden structures. A notable example is the NIST finalist Rainbow \cite{JD05}, a multivariate signature scheme that was broken in 2022 by Beullens \cite{Beu22} through exploitation of structural weaknesses.

		One might argue that post-quantum cryptography is not yet necessary, given that large-scale quantum computers capable of breaking classical cryptography do not currently exist. However, adversaries could perform so-called \emph{harvest-and-decrypt-later} attacks, in which encrypted data is collected today and decrypted in the future when quantum capabilities become available.

		In response to these concerns, the National Institute of Standards and Technology (NIST) launched an international competition in 2016 to standardize post-quantum cryptographic primitives.

		The \emph{MPC-in-the-Head} (MPCitH) paradigm, first introduced in \cite{IKOS07}, has recently gained significant attention—particularly in the second round of NIST’s post-quantum signature standardization process, where 5 out of 14 candidate schemes were based on this approach.

@@

		The main appeal of the MPCitH paradigm lies in the fact that it enables the construction of efficient, non-interactive, and quantum-resistant signature schemes with relatively small signature sizes and low verification cost and extremely small keys size.

@@

		The objective of this internship is to conduct a cryptanalysis of this scheme, assess its security and efficiency, and explore potential improvements.
		This includes evaluating its concrete security parameters and identifying potential structural weaknesses.

		In Section~\ref{sec2}, we introduce the MPCitH paradigm and provide an overview of the SBC problem, including its variants as well as the identification and signature schemes built upon it.

		In Section~\ref{sec3}, we present key results in polynomial system solving and describe the main algorithms and results used in our cryptanalysis.

		Finally, in Section~\ref{sec4}, we detail our contributions, including the cryptanalysis and formal security evaluation of the SBC-based signature scheme, and summarize the most efficient known attacks against it.

	\newpage
	\tableofcontents
	\newpage
	\section{MPCitH and SBC signature scheme}\label{sec2}

@@

		\subsubsection{GGM tree}
		Since most of the values used in the MPC protocol are randomly generated, a technique known as the GGM tree can be employed to transmit only a seed, rather than all the individual random values. This technique has been introduced in \cite{GGM86}.

@@

		Usually, the linear private share of a value $x$ is done this way:
		\begin{enumerate}
			\item we create $n$ random shares $x^{\lir} \overset{{\scriptscriptstyle\$}}\gets \mathbb{F}_q, \forall i \in [1,N]$.
			\item we compute the offset $\delta_x = x - \sum_{i=1}^{N} x^{\lir} \text{ mod } \mathbb{F}_q$.
		\end{enumerate}

		One challenge in this approach is the need to communicate the offset to the verifier so that they can verify the computations. However, a recent technique called \emph{Correlated GGM} (cGGM) has been introduced in \cite{GYWZ+22}, which allows the prover to generate random values $x^{\lir}$ using a modified GGM tree, such that $\sum_{i=1}^{N} x^{\lir} = x$, with minimal computational overhead.

@@

	\section{Polynomial system solving}\label{sec3}
		Gröbner basis theory was first introduced by Buchberger in his thesis \cite{Buc} and he gave an algorithm to compute the Gröbner basis of an ideal.
		Gröbner basis algorithms like Buchberger's Algorithm \cite{Buc}, are used to find solutions for polynomial systems using algebraic geometry tools. For a gentle introduction to this subject \cite{CLS}.

@@

	\section{Contributions}\label{sec4}

@@

		\begin{tcolorbox}[title=Protocol 1: Signature scheme - Signing Algorithm, breakable, enhanced, colback=white, colframe=black]\label{Protocol_sig}
			\textbf{Inputs:} A public key $(\vec{u}, \vec{v})$, a private key $(\vec{x}, \vec{y}) \in \text{NSBC}[\vec{u}, \vec{v}]$ and a message $m \in \{0,1\}^*$.

@@

			\vspace{0.5em}
			\textbf{Phase 4: Second challenge (parties to be opened).}
			\begin{enumerate}[label=\arabic*.]
				\item Compute $h := \text{Hash}(m, \texttt{salt}, \texttt{comm}^{[1]}, \dots, \texttt{comm}^{[\tau]})$.
				\item Expand $h$ as $(I^{[e]})_{e \in [1:\tau]} \leftarrow \text{Expand}(h)$ where, for every $e, I^{[e]} \in [1:N]$ the party that will not be opened.
			\end{enumerate}

@@

			$$
			\sigma := \texttt{salt} \parallel
			h \parallel
			punctured\_keys \parallel
			[\delta_A^{[e]}]_{e\in[1,\tau]} \parallel
			[\delta_B^{[e]}]_{e\in[1,\tau]} \parallel
			$$
			$$
			[\vec{\delta}_x^{[e]}]_{e\in[1,\tau]} \parallel [\vec{\delta}_y^{[e]}]_{e\in[1,\tau]} \parallel
			[(X_1 + t_0(\vec{u} \cdot \vec{x}))^{[e]}]_{e\in[1,\tau]} \parallel
			[(X_2 + t_0(\vec{v} \cdot \vec{x}))^{[e]}]_{e\in[1,\tau]} \parallel
			$$
			$$
			[(Y_1 + t_0(\vec{v} \cdot \vec{y}))^{[e]}]_{e\in[1,\tau]} \parallel
			[(Y_2 + t_0(\vec{u} \cdot \vec{y}))^{[e]}]_{e\in[1,\tau]}
			$$
		\end{tcolorbox}

		\begin{tcolorbox}[title=Protocol 2: Signature scheme - Verification Algorithm, breakable, enhanced, colback=white, colframe=black]\label{Protocol_verif}
			\textbf{Inputs:} A public key $(\vec{u}, \vec{v})$, a message $m \in \{0,1\}^*$ and a signature $\sigma$.

			\begin{enumerate} 
			\item Parse the signature $\sigma$ as
			$$
			\texttt{salt} \parallel
			h \parallel
			punctured\_keys \parallel
			[\delta_A^{[e]}]_{e\in[1,\tau]} \parallel
			[\delta_B^{[e]}]_{e\in[1,\tau]} \parallel
			$$
			$$
			[\vec{\delta}_x^{[e]}]_{e\in[1,\tau]} \parallel [\vec{\delta}_y^{[e]}]_{e\in[1,\tau]} \parallel
			[(X_1 + t_0(\vec{u} \cdot \vec{x}))^{[e]}]_{e\in[1,\tau]} \parallel
			[(X_2 + t_0(\vec{v} \cdot \vec{x}))^{[e]}]_{e\in[1,\tau]} \parallel
			$$
			$$
			[(Y_1 + t_0(\vec{v} \cdot \vec{y}))^{[e]}]_{e\in[1,\tau]} \parallel
			[(Y_2 + t_0(\vec{u} \cdot \vec{y}))^{[e]}]_{e\in[1,\tau]}
			$$

			\item Expand $h$ as $(I^{[e]})_{e \in [1:\tau]} \leftarrow \text{Expand}(h)$ where, for every $e, I^{[e]} \in [1:N]$ the party that we do not have.
			\end{enumerate}
		\end{tcolorbox}

@@

	This type of proof typically proceeds by simulating a game in which an adversary interacts with a signing oracle that responds to arbitrary signing queries. The adversary's goal is to produce a valid signature on a message that was not previously queried to the oracle. A signature scheme is said to be \textbf{EUF-CMA} (Existential Unforgeability under Chosen Message Attack) if the probability that the adversary succeeds in forging such a signature is at most $2^{-\lambda}$, where $\lambda$ denotes the security parameter. This probability is referred to as the adversary's advantage.
